---
title: "20230221_Cyto_POS-INT_PCA_v1"
author: "ASM"
date: '2023-03-04'
output:
  pdf_document: default
  html_document:
    theme: cosmo
    highlight: haddock
---
## Script used for PCA of proteomic samples

```{r setup, include = FALSE}
#knitr::opts_chunk$set(echo = TRUE, tidy = TRUE) # tidy = TRUE prevents text from running off the side of the knitted page.
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
setwd("/Users/amandameyer/Box Sync/McMahon Lab/Dissertation/TR01/BirA-AdipoqCre/Cytoplasmic/TR01_Adipose_cytoplasmic/") # git project linked to this folder

```


```{r libraries needed}
library(dplyr)
library(tidyr)
library(tidyverse)
library(ggvenn)
library(ggplot2)
library(preprocessCore)
library(qsmooth)
library(ggfortify)
library(devtools)
library(factoextra)
```




## Principal Component Analysis 

Analysis of Adipoq-BirA-G3 posterior subcutaneous (SAT) and interscapular brown (BAT) adipose tissue **proteomics** data to investiagte proteomic differences between SAT and BAT adipocytes using the **BirA-G3 cytoplasmic proximity labeling mouse**. ASM project history and version control can be found at:  <https://github.com/TaylorLS2020/TR01_Adipose_cytoplasmic>. 

Specifically, this analysis is using LC-MS/MS (MS3) intensity data from UCLA.  


```{r PCA}
# read in intensity data 
cyto.int <- read.csv("original_data_UCLA/2022-10-23-Amanda-Meyer-Andrew-McMahon-SET-A-proteinGroups.csv")

# impute NAs
cyto.int[is.na(cyto.int)] <- 10

# matrix for quaterile quant
cyto.mtx <- as.matrix(cyto.int[ , 25:34])

# quartiles
cyto.qt <- normalize.quantiles(cyto.mtx)

# qsmooth normalization
cyto.sm <- qsmooth(cyto.mtx, group_factor = c(0, 0, 1, 1, 1, 1, 2, 2, 2, 2))
qsmoothWeights(cyto.sm)

# lable dfs 
cytoCols <- c("CTR01", "CTR02", "POS01", "POS02", "POS03", "POS04", "INT01", "INT02", "INT03", "INT04")
cytoCond <- c("CTR", "CTR", "POS", "POS", "POS", "POS", "INT", "INT", "INT", "INT")
colnames(cyto.qt) <- cytoCols

# data frame setup for pca 
cytoq.df <- as.data.frame(t(cyto.sm@qsmoothData[ , ]))
cytoq.df$group <- cytoCols 
cytoq.df$Condition <- cytoCond
cytoq.df$Sex <- c("M", "M", "F", "F", "M", "M", 
                  "F", "F", "M", "M")
cytoq.df$Tissue <- c("POS", "INT", "POS", "POS", "POS", "POS", 
                     "INT", "INT", "INT", "INT")

```



# PCA by Tissue and Genotype
```{r PCA Plot by Tissue and Genotype, echo=T, fig.cap='PCA of Adipoq-BirA-G3 and BirA-G3 from SAT and BAT'}
# PCA using prcomp and plot 
res.pca <- prcomp(cytoq.df[ , 1:1534],  scale. = T) # change range to exclude group/condition
(plot1<- autoplot(res.pca, data = cytoq.df, colour = "Condition", alpha = 0.85, size = 2.5) + 
    theme_classic() + 
    scale_color_manual(labels = c("CTR", "SAT", "BAT"), 
                       values = c("grey70", "#D53E4F", "#6BAED6")))

ggsave(filename = "Cyto_POS_INT_LCMSMS_PCA_v1.pdf", plot = plot1, path = "figures/", 
       width = 5, height = 4, units = "in", dpi = "retina")

```



# PCA by Tissue, Genotype, and Sex
```{r PCA Plot by Tissue and Sex, echo=T, fig.cap='PCA of Adipoq-BirA-G3 and BirA-G3 from male and female SAT and BAT'}
# labeled by sex
(plot1<- autoplot(res.pca, data = cytoq.df, colour = "Condition", shape = "Sex", alpha = 0.80, size = 2.5) + 
    theme_classic() + 
    scale_color_manual(labels = c("CTR", "SAT", "BAT"), 
                       values = c("grey70","#6BAED6", "#42853D")))


ggsave(filename = "Cyto_POS_INT_wSexlabels_LCMSMS_PCA_v1.pdf", plot = plot1, path = "figures/", 
       width = 5, height = 4, units = "in", dpi = "retina")

```


```{r session}
sessionInfo()

```


